---
title: "viz_1"
author: "Kimia Faroughi"
date: "2025-09-25"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
```

Import the weather data.

```{r}
data("weather_df")
```

Making our first plot :-)

```{r}
ggplot(data = weather_df, mapping = aes(x = tmin, y = tmax)) + 
  geom_point() #ppl never write data = or mapping = arguments

ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
weather_df |> #using pipe operator instead
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggp_weather_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_weather_scatterplot #prints the plot
```

Check that some rows are missing.

```{r}
weather_df |> 
  filter(is.na(tmax)) #lists na obs for tmax
```

## Fancier scatterplots!

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + #gives different color for each name variable in dataset
  geom_point(alpha = 0.3) + #makes points a little see through
  geom_smooth(se = FALSE) #adds smooth lines for each subcategory and se = FALSE removes suspect error band (don't trust them)
```

Where you define aesthetics can matter.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, size = prcp), alpha = 0.3) + #changes size of every point based on precipitation var
  geom_smooth(se = FALSE)

#produces only one smooth line (not divided by subcategory) because moved the aesthetic of color to geom_point rather than the main ggplot line
#communicate one message at a time with one or two aesthetic mappings, don't need to include lots of different options because you want your "audience" to understand the plot
```

Use faceting. 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) #separates into different plots based on variable name (by column)

weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.3, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .) #separates into different plots based on variable name (by row)
```

Let's make a somewhat more interesting scatterplot.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + #shape = name creates different shapes as the points based on variable name
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Learning Assessment

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_F = (tmax * (9/5)) + 32,
    tmin_F = (tmin * (9/5)) + 32
  ) |> 
  ggplot(aes(x = tmin_F, y = tmax_F)) +
  geom_point() +
  geom_smooth(se = FALSE, method = "lm") #adds linear regression line
```

## Small things

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  #geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) 
#don't need to have points on the plot

weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) +
  geom_smooth(se = FALSE) +
  geom_point()
#paints points on plot after the smooth, so order matters
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() #can be helpful in avoiding overplotting
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "#086bfd")
```

## Univariate plots

```{r}
#histogram
weather_df |> 
  ggplot(aes(x = tmin)) +
  geom_histogram(color = "white", fill = "red")

weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() #overlays when fill = name which is confusing
```

How to fix this.

```{r}
#facet grid
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .)

#density plot
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.2)

#boxplot
weather_df |> 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot(aes(fill = name))

#violin plot
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()

#ridge plot
weather_df |> 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

Learning assessment for univariate plots.

```{r}
#histogram
weather_df |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram() +
  facet_grid(name ~ .)

#density plot
weather_df |> 
  filter(prcp > 5, prcp < 1000) |> #zooms in into data, need to communicate that only part of it is being analyzed when you move forward
  ggplot(aes(x = prcp, fill = name)) +
  geom_density() 

#boxplot
weather_df |> 
  ggplot(aes(x = name, y = prcp)) +
  geom_boxplot(aes(fill = name))

#violin plot
weather_df |> 
  ggplot(aes(x = name, y = prcp, fill = name)) +
  geom_violin()

#ridge plot
weather_df |> 
  ggplot(aes(x = prcp, y = name)) +
  geom_density_ridges()
```

## Saving and embedding plots

Saving plots.

```{r}
ggp_weather_violin = 
  weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()

ggp_weather_violin

ggsave("plots/violin_plot.pdf", ggp_weather_violin, 
       width = 8, height = 6) #width and height in inches
```

Embedding plots.

```{r, fig.width = 6, fig.eight = 4}
ggp_weather_violin
```